Comparison of Confidence Interval in the OLS and Ridge Linear Regression Model: A Comparative study via Simulation

### 2024-08-28 draft_1.qmd
introduction edit

Multiple linear regression maps the relationship between two or more predictors and dependent variable to a linear equation. The aim is to predict the response variable uisng the indepnedent variables. If X is a n×p full rank matrix of predictors and Y is a n × 1 vector of response variables the multiple linear regression can be explained as,

$$Y = X \times \beta + \epsilon$$,
where, $\beta$ is an p × 1 unknown regression paramaters and $\epsilon$ is the n × 1 vector of error with mean zero and equal variance.

One of the key assumption of the widely used multiple linear regression model is that the predictors need to be independent of each other. Violating the assumption of independence results in an issue known as multicollinearity. Multicollinearity occurs when two or more independent variables are highly correlated with each other. It causes standard error of the coefficient to increase resulting in wide confidence interval and less reliable results. Identifying significant predictors also becomes difficult.

@@

@@

>>>>

Multiple linear regression maps the relationship between two or more predictors and dependent variable to a linear equation. The aim is to predict the response variable using the independent variables. If X is a n×p full rank matrix of predictors and Y is a n × 1 vector of response variables the multiple linear regression can be explained as,

$$Y = X \times \beta + \epsilon, $$
where, $\beta$ is an p × 1 unknown regression paramaters and $\epsilon$ is the n × 1 vector of error with mean zero and equal variance.Ordinary Least Square (OLS) method is commonly used to estimate the unknown regression parameter. The OLS estimate in case of linear regression model is defined as follows,

$$\hat{\beta} = (X^T X)^{-1} X^T y.$$

One of the key assumption of the widely used multiple linear regression model is that the predictors need to be independent of each other. Violating the assumption of independence 
results in an issue known as multicollinearity. Multicollinearity occurs when two or more independent variables are highly correlated with each other. It causes standard error of the OLS coefficient to increase resulting in wide confidence intervals and less reliable results. Identifying and interpreting significant predictors also becomes difficult. As the field of study progressed, researchers developed a number of methods to address presence of multicollinearity in data. Ridge regression [@hoerl1970ridge], Lasso,

Ridge regression is one of the methods proposed by @hoerl1970ridge that mitigates the effects of multicollinearity by introducing an amount of bias in to the model.

para about differnt techniques,

para about ridge

para about CI

para about Aim

Note:

@@

## CI for Ridge regression

@@

# Bibliography

### 2024-08-28 draft_1.tex
introduction edit

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

@@

One of the key assumption of the widely used multiple linear regression
model is that the predictors need to be independent of each other.
Violating the assumption of independence results in an issue known as
multicollinearity. Linear regression maps the relationship between the
predictors and dependent variable with a linear equation. If X is a n×p
full rank matrix of predictors and Y is a n × 1 vector of response
variables the multiple linear regression can be explained as,

\[Y = X \times \beta + \epsilon\], where, \(\beta\) is an p × 1 unknown
regression paramaters and \[\epsilon\] is the n × 1 vector of error with
mean zero and equal variance.\\
\# Statistical Methodology

@@

@@

>>>>

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

Multiple linear regression maps the relationship between two or more
predictors and dependent variable to a linear equation. The aim is to
predict the response variable using the independent variables. If X is a
n×p full rank matrix of predictors and Y is a n × 1 vector of response
variables the multiple linear regression can be explained as,

\[Y = X \times \beta + \epsilon, \] where, \(\beta\) is an p × 1 unknown
regression paramaters and \(\epsilon\) is the n × 1 vector of error with
mean zero and equal variance.Ordinary Least Square (OLS) method is
commonly used to estimate the unknown regression parameter. The OLS
estimate in case of linear regression model is defined as follows,

\[\hat{\beta} = (X^T X)^{-1} X^T y.\]

One of the key assumption of the widely used multiple linear regression
model is that the predictors need to be independent of each other.
Violating the assumption of independence results in an issue known as
multicollinearity. Multicollinearity occurs when two or more independent
variables are highly correlated with each other. It causes standard
error of the OLS coefficient to increase resulting in wide confidence
intervals and less reliable results. Identifying and interpreting
significant predictors also becomes difficult. As the field of study
progressed, researchers developed a number of methods to address
presence of multicollinearity in data. Ridge regression (Hoerl and
Kennard 1970), Lasso,

Ridge regression is one of the methods proposed by Hoerl and Kennard
(1970) that mitigates the effects of multicollinearity by introducing an
amount of bias in to the model.

para about differnt techniques,

para about ridge

para about CI

para about Aim

Note:

\section{Statistical Methodology}\label{statistical-methodology}

@@

\subsection{Confidence Interval}\label{confidence-interval}

\subsection{CI for Ridge regression}\label{ci-for-ridge-regression}

@@

\section*{Bibliography}\label{bibliography}
\addcontentsline{toc}{section}{Bibliography}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-hoerl1970ridge}
Hoerl, Arthur E, and Robert W Kennard. 1970. {``Ridge Regression: Biased
Estimation for Nonorthogonal Problems.''} \emph{Technometrics} 12 (1):
55--67.

\end{CSLReferences}
